The Rockerverse: Images, Packages, and Applications for Containerization with R

### 2020-01-20 manuscript.Rmd
small text update
Co-authored-by: Dirk Eddelbuettel <edd@debian.org>

## Non-Debian Linux images
\label{nondebian}

@@

_Data Science_ is a widely discussed topic among all academic disciplines \citep[e.g.,][]{donoho_50_2017}.
Using computers and complex software to make sense of (large amounts of) collected or simulated data is a cross-cutting phenomenon leading to _X Data Science_ disciplines, seminars, and study majors.
The practice to apply a combination tools and software stacks for analysing data and the cross-cutting skillsets connected with data science puts the management of computing environments and the challenges of reproducibility high up on the agenda.
This awareness leads to a large number of bespoke Docker images but also some commonly used image stacks, ranging from base images with very common tools to 'catch all' images including _a lot_ of software.
Here we only mention a few image stacks as examples, which are not based on Rocker images or stem from the R community, but prominently include installations of R.
A general description of what to consider when creating a Docker image used for data science projects in R is provided on [environments.rstudio.com](https://environments.rstudio.com/docker).

@@

**Kaggle** provides the [`gcr.io/kaggle-images/rstats`](https://hub.docker.com/r/kaggle/rstats) image (previously `kaggle/rstats`) and [corresponding `Dockerfile`](https://github.com/Kaggle/docker-rstats) for usage in their Machine Learning competitions and easy access to the associated datasets.
It includes machine learling libraries such as Tensorflow and Keras, and also configures the \CRANpkg{reticulate} package.
The image uses a base image with _all packages from CRAN_, `gcr.io/kaggle-images/rcran`, which requires a Google Cloud Build as Docker&nbsp;Hub would time out\footnote{Originally, a stacked collection of over 20 images with automated builds on Docker~Hub was used, see \href{https://web.archive.org/web/20190606043353/http://blog.kaggle.com/2016/02/05/how-to-get-started-with-data-science-in-containers/}{https://web.archive.org/web/20190606043353/http://blog.kaggle.com/2016/02/05/how-to-get-started-with-data-science-in-containers/} and \href{https://hub.docker.com/r/kaggle/rcran/dockerfile}{https://hub.docker.com/r/kaggle/rcran/dockerfile}}.
The final extracted image size is over 25GB, which makes it debatable if having everything available is actually convenient.

@@

Using **GPUs** (graphical processing units) as a specialised hardware from containerized common work environments is also possible and useful \citep{haydel_enhancing_2015}.
GPUs are increasingly popular for compute-intensive machine learning (ML) tasks, e.g., deep artificial neural networks \citep{schmidhuber_deep_2015}.
Though in this case, containers are not completely portable between hardware environments, but the software stack for ML with GPUs is so complex to set up that a ready-to-use container is helpful.
Containers running GPU software require drivers and libraries specific to GPU models and versions, and containers require a specialized runtime to connect to the underlying GPU hardware.
For NVIDIA GPUs, the [NVIDIA Container Toolkit](https://github.com/NVIDIA/nvidia-docker) includes a specialized runtime plugin for Docker and a set of base images with appropriate drivers and libraries.
The Rocker&nbsp;Project [has a repository](https://github.com/rocker-org/ml) with (beta) images based on these that include GPU-enabled versions of machine-learning R packages, e.g., `rocker/ml` and `rocker/tensorflow-gpu`.

>>>>

## Non-Rocker Linux images
\label{nondebian}

@@

_Data Science_ is a widely discussed topic among all academic disciplines \citep[e.g.,][]{donoho_50_2017}.
The discussions shed a light on the tools and craftspersonship behind the analysis of data with computational methods.
The practice of Data Science often involves a combination tools and software stacks and requires a cross-cutting skillset.
This complexity and an inherent concern for openness and reproducibility in the Data Science community lead to Docker being used widely.
<!--A general description of what to consider when creating a Docker image used for data science projects in R is provided on [environments.rstudio.com](https://environments.rstudio.com/docker).-->
This section presents exemplary Docker images and image stacks featuring R intended for Data Science.

@@

**Kaggle** provides the [`gcr.io/kaggle-images/rstats`](https://hub.docker.com/r/kaggle/rstats) image (previously `kaggle/rstats`) and [corresponding `Dockerfile`](https://github.com/Kaggle/docker-rstats) for usage in their Machine Learning competitions and easy access to the associated datasets.
It includes machine learning libraries such as Tensorflow and Keras (see also image `rocker/ml` in Section&nbsp;\nameref{rocker-gpu}), and also configures the \CRANpkg{reticulate} package.
The image uses a base image with _all packages from CRAN_, `gcr.io/kaggle-images/rcran`, which requires a Google Cloud Build as Docker&nbsp;Hub would time out\footnote{Originally, a stacked collection of over 20 images with automated builds on Docker~Hub was used, see \href{https://web.archive.org/web/20190606043353/http://blog.kaggle.com/2016/02/05/how-to-get-started-with-data-science-in-containers/}{https://web.archive.org/web/20190606043353/http://blog.kaggle.com/2016/02/05/how-to-get-started-with-data-science-in-containers/} and \href{https://hub.docker.com/r/kaggle/rcran/dockerfile}{https://hub.docker.com/r/kaggle/rcran/dockerfile}}.
The final extracted image size is over 25GB, which makes it debatable if having everything available is actually convenient.

@@

\label{rocker-gpu}
Using **GPUs** (graphical processing units) as a specialised hardware from containerized common work environments is also possible and useful \citep{haydel_enhancing_2015}.
GPUs are increasingly popular for compute-intensive machine learning (ML) tasks, e.g., deep artificial neural networks \citep{schmidhuber_deep_2015}.
Though in this case, containers are not completely portable between hardware environments, but the software stack for ML with GPUs is so complex to set up that a ready-to-use container is helpful.
Containers running GPU software require drivers and libraries specific to GPU models and versions, and containers require a specialized runtime to connect to the underlying GPU hardware.
For NVIDIA GPUs, the [NVIDIA Container Toolkit](https://github.com/NVIDIA/nvidia-docker) includes a specialized runtime plugin for Docker and a set of base images with appropriate drivers and libraries.
The Rocker&nbsp;Project [has a repository](https://github.com/rocker-org/ml) with (beta) images based on these that include GPU-enabled versions of machine-learning R packages, e.g., `rocker/ml` and `rocker/tensorflow-gpu`.
